
---
title: "Rene Guerra - Homework 3"
output: pdf_document
date: 'Sunday, October 30th 2022'
editor_options: 
  markdown: 
    wrap: 72
---

```{r include= FALSE}

library(tidyverse)
library(tidymodels)
library(ggplot2)
library(readr)
library(corrr)
library(discrim)
library(klaR)
library(pROC)
tidymodels_prefer()

```

```{r include= FALSE}

Titanic <- read_csv("Titanic.csv")

```

```{r}

Titanic$survived <- factor(Titanic$survived, levels= c('Yes', 'No'))
Titanic$pclass <- factor(Titanic$pclass)

levels(Titanic$survived)

```

**1**

```{r}

set.seed(2022)

Titanic_split <- initial_split(Titanic, prop= 0.80, strata= survived)

Titanic_train <- training(Titanic_split)
Titanic_train

Titanic_test <- testing(Titanic_split)
Titanic_test

```

It is a good idea to use stratified sampling for this data because the sample comes from all categories and is divided into subcategories that will potentially derive different results for the outcome variable we are evaluating.


**2**

```{r}

Titanic_train %>%
  ggplot(aes(x= survived)) + geom_bar()

```

Most people did not survive based on the training data set.


**3**

```{r}

is.numeric(Titanic$passenger_id)
is.numeric(Titanic$survived)
is.numeric(Titanic$pclass)
is.numeric(Titanic$name)
is.numeric(Titanic$sex)
is.numeric(Titanic$age)
is.numeric(Titanic$sib_sp)
is.numeric(Titanic$parch)
is.numeric(Titanic$ticket)
is.numeric(Titanic$fare)
is.numeric(Titanic$cabin)
is.numeric(Titanic$embarked)

```


```{r}

Cor_Titanic <- Titanic_train %>%
  select(-c(survived, pclass, name, sex, ticket, cabin, embarked)) %>%
  correlate()
rplot(Cor_Titanic)

Cor_Titanic %>%
  stretch() %>%
  ggplot(aes(x, y, fill= r)) +
  geom_tile() +
  geom_text(aes(label= as.character(fashion(r))))


```
The main pattern I see is the negative correlation between sib_sp and age.
The correlation matrix shows that older passengers were not accompanied by 
siblings and younger passengers did not have spouses.
Another significant negative correlation is parch and age. Older passengers
did not have parents on board and younger passengers were not accompanied by 
children. On the other hand, sib_sp and parch have a significantly positive
correlation.
This means that most children and siblings had parents aboard,
thus there were families.


**4**

```{r}

Titanic_recipe <- recipe(survived ~ pclass + sex + age + sib_sp + parch + fare, Titanic_train) %>%
  step_impute_linear(age, impute_with = imp_vars(sib_sp)) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_interact(~ starts_with("sex"):fare + age:fare)

Titanic_recipe

```

```{r warning= FALSE}

Interaction1 <- lm(survived ~ sex + fare, data= Titanic_train)
Interaction1

Interaction2 <- lm(survived ~ age + fare, data= Titanic_train)
Interaction2

```

**5**

```{r}

glm_Titanic <- logistic_reg() %>%
  set_engine(("glm")) %>%
  set_mode("classification")
  
glm_Titanic

```

```{r}

glm_Titanicflow<- workflow() %>%
  add_model(glm_Titanic) %>%
  add_recipe(Titanic_recipe)

TitanicFit1 <- fit(glm_Titanicflow, Titanic_train)

TitanicFit1 %>%
  tidy()

predict(TitanicFit1, new_data = Titanic_train, type= "prob")

augment(TitanicFit1, new_data= Titanic_train) %>%
  conf_mat(truth= survived, estimate= .pred_class) %>%
  autoplot(type= "heatmap")

```

```{r}

glm_accuracy <- augment(TitanicFit1, new_data= Titanic_train) %>%
  accuracy(truth= survived, estimate= .pred_class)

glm_accuracy

```

**6**

```{r}

lda_Titanic <- discrim_linear() %>% 
  set_engine("MASS") %>%
  set_mode("classification")

lda_Titanic

```

```{r}

lda_TitanicFlow <- workflow() %>% 
  add_model(lda_Titanic) %>% 
  add_recipe(Titanic_recipe)

TitanicFit2 <- fit(lda_TitanicFlow, Titanic_train)

predict(TitanicFit2, new_data = Titanic_train, type= "prob")

augment(TitanicFit2, new_data= Titanic_train) %>%
  conf_mat(truth= survived, estimate= .pred_class) %>%
  autoplot(type= "heatmap")

```

```{r}

lda_accuracy <- augment(TitanicFit2, new_data= Titanic_train) %>%
  accuracy(truth= survived, estimate= .pred_class)

lda_accuracy

```

**7**

```{r}

qda_Titanic <- discrim_quad() %>% 
  set_engine("MASS") %>%
  set_mode("classification")

qda_Titanic

```

```{r}

qda_TitanicFlow <- workflow() %>% 
  add_model(qda_Titanic) %>% 
  add_recipe(Titanic_recipe)

TitanicFit3 <- fit(qda_TitanicFlow, Titanic_train)

predict(TitanicFit3, new_data = Titanic_train, type= "prob")

augment(TitanicFit3, new_data= Titanic_train) %>%
  conf_mat(truth= survived, estimate= .pred_class) %>%
  autoplot(type= "heatmap")

```

```{r}

qda_accuracy <- augment(TitanicFit3, new_data= Titanic_train) %>%
  accuracy(truth= survived, estimate= .pred_class)

glm_accuracy

```

**8**

```{r}

nB_Titanic <- naive_Bayes() %>% 
  set_engine("klaR") %>%
  set_mode("classification") %>%
  set_args(usekernel= FALSE)

nB_Titanic

```

```{r warning= FALSE}

nB_TitanicFlow <- workflow() %>% 
  add_model(nB_Titanic) %>% 
  add_recipe(Titanic_recipe)

TitanicFit4 <- fit(nB_TitanicFlow, Titanic_train)

predict(TitanicFit4, new_data= Titanic_train, type= "prob")

augment(TitanicFit4, new_data= Titanic_train) %>%
  conf_mat(truth= survived, estimate= .pred_class) %>%
  autoplot(type= "heatmap")

```

```{r warning= FALSE}

nB_accuracy <- augment(TitanicFit4, new_data= Titanic_train) %>%
  accuracy(truth= survived, estimate= .pred_class)

nB_accuracy

```

**9**

```{r}

TotalAccurate <- c(glm_accuracy$.estimate, lda_accuracy$.estimate, qda_accuracy$.estimate, nB_accuracy$.estimate)

ModType <- c("Logistic Regression", "LDA", "QDA", "Naive Bayes")

Performance <- tibble(TotalAccurate= TotalAccurate, ModType= ModType)

Performance %>%
  arrange(-TotalAccurate)

```

The model that achieved the highest accuracy on the training data is
Logistic Regression with 0.8160112 accuracy.


**10**

```{r}

predict(TitanicFit1, new_data= Titanic_test, type= "prob")

augment(TitanicFit1, new_data= Titanic_test) %>%
  conf_mat(truth= survived, estimate= .pred_class) %>%
  autoplot(type= "heatmap")

```

```{r}

Add_metric <- metric_set(accuracy, sensitivity, specificity)

augment(TitanicFit1, new_data= Titanic_test) %>%
  Add_metric(truth= survived, estimate= .pred_class)

```

```{r}

augment(TitanicFit1, new_data= Titanic_test) %>%
  roc_curve(survived, .pred_No) %>%
  autoplot()


```
The ROC curve is below the random classifier.
Bad model for classification of
variable for no survival. The area under the curve is 0.87
The training accuracy is 0.8160112.
The testing accuracy is 0.7988827.
Accuracy results are not significantly different and it is normal to have a
higher training accuracy since the model is optimized to train data.
